---
title: "ExampleClustering"
author: "Maximilian Lombardo"
date: "8/15/2020"
output: html_document
params:
  raw.data.parent.dir: ""
  sample.size: 1000
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Necessary libraries

```{r}
library(Seurat)
```

Read in raw data and generate Seurat object

```{r}

raw.dirs <- list.dirs(params$raw.data.parent.dir, recursive = FALSE)

#Read in counts - we have multiple runs/batchs so we will need to comnbine these
#in a later step
counts.1 <- Read10X(raw.dirs[1])
counts.2 <- Read10X(raw.dirs[2])
counts.3 <- Read10X(raw.dirs[3])
counts.4 <- Read10X(raw.dirs[4])

#Sampling Cells from each count matrix for demonstration purposes
sampleCells <- function(raw.counts){
  sampled.counts <- raw.counts[, sample(x = colnames(raw.counts),
                                    size = params$sample.size,
                                    replace = FALSE)]
  return(sampled.counts)
}

#Sampling counts for this demo - some of these datasets are quite large
sampled.counts.1 <- sampleCells(counts.1)
sampled.counts.2 <- sampleCells(counts.2)
sampled.counts.3 <- sampleCells(counts.3)
sampled.counts.4 <- sampleCells(counts.4)


#Generate Seurat objects for individual runs
obj.1 <- CreateSeuratObject(sampled.counts.1, project = "Batch 1")
obj.2 <- CreateSeuratObject(sampled.counts.2, project = "Batch 2")
obj.3 <- CreateSeuratObject(sampled.counts.3, project = "Batch 3")
obj.4 <- CreateSeuratObject(sampled.counts.4, project = "Batch 4")

#Combine Seurat objects 
obj <- merge(obj.1, c(obj.2, obj.3, obj.4))

#Remove the original count matrices to free up some memory
#rm(counts.1, counts.2,
#   counts.3, counts.2)


```

Quality Control

```{r}
#Add mitochondrial fraction as a metadata item
obj[['percent.mito']] <- PercentageFeatureSet(obj, pattern = "^mt-")

#Visualize QC feature distributions
VlnPlot(obj,
        features = c("nFeature_RNA", "nCount_RNA", "percent.mito"),
        ncol = 3, pt.size = 0.1)
```
Subset cells with low mitohcondrial fraction

```{r}
obj <- subset(obj, percent.mito < 10)

```


Standard Normalization, Variable Feature Selection, and Scaling

```{r}
obj <- NormalizeData(obj)
obj <- FindVariableFeatures(obj)
obj <- ScaleData(obj, features = rownames(obj[["RNA"]]@data))

```

Linear Dimensional reduction

```{r}
obj <- RunPCA(obj, features = VariableFeatures(object = obj))


#Visualize contributions of PCs
ElbowPlot(obj, ndims = 50)


#set number of dimensions for
#downstream clustering and visualization
ndims <- 1:20
```


Visualization

```{r}
obj <- RunUMAP(obj, dims = ndims,
               n.neighbors = 10,
               local.connectivity = 10)

obj <- RunTSNE(obj, dims = ndims, perplexity = 30, max_iter = 2000, theta = 0)
```

View our Data in 2d

```{r}
TSNEPlot(obj)
UMAPPlot(obj)
```

Find Clusters and corresponding markers

```{r}
obj <- FindNeighbors(obj, dims = ndims, k.param = 10)

obj <- FindClusters(obj, resolution = 0.8)

obj@misc$markers <- FindAllMarkers(obj, min.pct = 0.5)

```

Save atlas Object

```{r}
#Need to chage for your machine
saveRDS(obj, "~/Documents/DemoLocation/objectName.rds")
```


Subset the object without the 4th batch and filter out the cells with
a high mitochondrial fraction

```{r}
healthy.cells <- rownames(obj@meta.data[!obj@meta.data$orig.ident %in% 'Batch 4',])

obj.sub <- subset(obj, cells = healthy.cells)

#This is pretty stringent, but we only want the high quality data
obj.sub <- subset(obj.sub, percent.mito < 10)

```

Standard Normalization, Variable Feature Selection, and Scaling on the 
object subset

```{r}
obj.sub <- NormalizeData(obj.sub)
obj.sub <- FindVariableFeatures(obj.sub)
obj.sub <- ScaleData(obj.sub, features = rownames(obj.sub[["RNA"]]@data))
```

Visualization

```{r}
obj.sub <- RunUMAP(obj.sub, dims = ndims,
               n.neighbors = 10,
               local.connectivity = 10)

obj.sub <- RunTSNE(obj.sub, dims = ndims,
                   perplexity = 30,
                   max_iter = 2000, theta = 0)
```

View our object in 2d

```{r}
TSNEPlot(obj.sub, group.by = 'orig.ident')
UMAPPlot(obj.sub, group.by = 'orig.ident')
```

Find Clusters and corresponding markers

```{r}
obj.sub <- FindNeighbors(obj.sub, dims = ndims, k.param = 10)

obj.sub <- FindClusters(obj.sub, resolution = 0.8)

obj.sub@misc$markers <- FindAllMarkers(obj.sub, min.pct = 0.5)

```

Save the subsetted data

```{r}
saveRDS(obj.sub, "~/Documents/DemoLocation/objectNameSubset.rds")
```
